---
title: "Much ado about fitness correlations"
author: "Megan Frederickson"
date: `r format(Sys.Date())`
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Fitness conflict or fitness alignment?

This repository re-analyzes the data in: Gano-Cohen KA, Wendlandt CE, Stokes PJ, Blanton MA, Quides KW, Zomorrodian A, Adinata ES, Sachs JL (2019) Interspecific conflict and the evolution of ineffective rhizobia. Ecology Letters. https://doi.org/10.1111/ele.13247

First we need to load some nifty R packages.

```{r Load packages}

#Load useful packages
library(tidyverse) #includes ggplot2, dplyr, readr, stringr
library(cowplot)
library(car)
library(glmm)
library(nlme)
library(lme4)
library(readxl)

```

I downloaded the data from Dryad on April 16, 2019. The citation for the data package is:

Gano-Cohen KA, Wendlandt CE, Stokes PJ, Blanton MA, Quides KW, Zomorrodian A, Adinata ES, Sachs JL (2019) Data from: Interspecific conflict and the evolution of ineffective rhizobia. Dryad Digital Repository. https://doi.org/10.5061/dryad.cr65269

```{r Download data}

data <- read_csv("Table_S4.csv")

```

