---
title: "Much ado about fitness correlations"
author: "Megan Frederickson"
date: `r format(Sys.Date())`
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Fitness conflict or fitness alignment?

This repository re-analyzes the data in: Gano-Cohen KA, Wendlandt CE, Stokes PJ, Blanton MA, Quides KW, Zomorrodian A, Adinata ES, Sachs JL (2019) Interspecific conflict and the evolution of ineffective rhizobia. Ecology Letters. https://doi.org/10.1111/ele.13247

First we need to load some nifty R packages.

```{r Load packages}

#Load useful packages
library(tidyverse) #includes ggplot2, dplyr, readr, stringr
library(cowplot)
library(car)
library(glmm)
library(nlme)
library(lme4)
library(emmeans)

```

I downloaded the data from Dryad on April 16, 2019. The citation for the data package is:

Gano-Cohen KA, Wendlandt CE, Stokes PJ, Blanton MA, Quides KW, Zomorrodian A, Adinata ES, Sachs JL (2019) Data from: Interspecific conflict and the evolution of ineffective rhizobia. Dryad Digital Repository. https://doi.org/10.5061/dryad.cr65269

```{r Download data}

data <- read_csv("Table_S4.csv")

```

The paper has several measures of plant and rhizobium fitness. First, we need to calculate the appropriate strain means (i.e., breeding values) for each variable. I have tried to stick as close as possible to the analysis presented in the paper, which states that "all effects were coded as fixed except block, which was treated as a random effect" and that variables were log-transformed to improve normality. I calculated breeding values on sympatric and universal hosts, again following the paper's general modelling approach.

```{r Genetic values}

#Exclude control, uninoculated plants

data <- subset(data, Strain != "control")

#Exclude plants that did not form nodules

data <- subset(data, `Total nodules` > 0)

#Make sure factors are factors and numbers are numbers

data$Strain <- as.factor(data$Strain)
data$`Host Line` <- as.factor(data$`Host Line`)
data$Population <- as.factor(data$Population)
data$`Total nodules` <- as.numeric(data$`Total nodules`)
data$`Mean individual  nodule biomass (mg)` <- as.numeric(data$`Mean individual  nodule biomass (mg)`)
data$`Shoots mass (g)` <- as.numeric(data$`Shoots mass (g)`)
data$`Roots mass (g)` <- as.numeric(data$`Roots mass (g)`)
data$`Relative Growth` <- as.numeric(data$`Relative Growth`)

#Log transform as needed, as per paper

data$log_nodules <- log(data$`Total nodules`)
data$log_mean_nod_biomass <- log(data$`Mean individual  nodule biomass (mg)`)

#Combine shoot and root mass into one measure of plant biomass
data$plant_biomass <- data$`Shoots mass (g)` + data$`Roots mass (g)`

#Divide into sympatric and universal hosts, as per paper

data_sym <- subset(data, `Host Line` != "A. heermannii" & `Host Line` != "UnH: Cla12.04" & `Host Line` != "UnL: Anz13.04")
data_uni <- subset(data, `Host Line` == "A. heermannii" | `Host Line` == "UnH: Cla12.04" | `Host Line` == "UnL: Anz13.04")

#Model total nodules
lmm1 <- lmer(log_nodules~Population + Strain + `Host Line` + (1|Block), data=data_sym) 
plot(lmm1)
summary(lmm1)
Anova(lmm1, type=3)

#Extract line means and save them to new dataframe

df.means <- as.data.frame(emmeans(lmm1, "Strain"))
colnames(df.means) <- c("Strain", "Population", "tot_nod_lsmean", "tot_nod_SE", "tot_nod_df", "tot_nod_lowCL", "tot_nod_highCL")

#Mean nodule mass
lmm2 <- lmer(log_mean_nod_biomass~Population + Strain + `Host Line` + (1|Block), data=data_sym)
plot(lmm2)
summary(lmm2)
Anova(lmm2, type=3)

#Extract line means and add them to the dataframe

tmp <- as.data.frame(emmeans(lmm2, "Strain"))
colnames(tmp) <- c("Strain", "Population", "nod_mass_lsmean", "nod_mass_SE", "nod_mass_df", "nod_mass_lowCL", "nod_mass_highCL")
df.means <- cbind(df.means, tmp[,3:7])

#Plant biomass

lmm3 <- lmer(plant_biomass~Population + Strain  + `Host Line` + (1|Block), data=data_sym)
plot(lmm3)
summary(lmm3)
Anova(lmm3, type=3)

#Extract line means and add them to the dataframe

tmp <- as.data.frame(emmeans(lmm3, "Strain"))
colnames(tmp) <- c("Strain", "Population", "plant_biomass_lsmean", "plant_biomass_SE", "plant_biomass_df", "plant_biomass_lowCL", "plant_biomass_highCL")
df.means <- cbind(df.means, tmp[,3:7])

#Relative growth rate, as per paper
#The paper says they took the log of relative growth rate, but there are negative numbers, hmmm...

hist(data_sym$`Relative Growth`) 

lmm4 <- lmer(log(`Relative Growth`)~Population + Strain  + `Host Line` + (1|Block), data=data_sym)
plot(lmm4)
summary(lmm4)
Anova(lmm4, type=3)

#Extract line means and add them to the dataframe

tmp <- as.data.frame(emmeans(lmm4, "Strain"))
colnames(tmp) <- c("Strain", "Population", "RGR_lsmean", "RGR_SE", "RGR_df", "RGR_lowCL", "RGR_highCL")
df.means <- cbind(df.means, tmp[,3:7])

#Add CHR and SI values for each isolate; they are identical for each isolate, so no need to model them

tmp <- data_sym %>% group_by(Strain) %>% summarize(CHR = mean(`CHR local abundance`), SI = mean(`SI local abunance`))
df.means <- merge(df.means, tmp, by="Strain")

#I can infer the biomass mass of plants in the controls for each strain just by simple math
#If RGR is inoculated biomass/control biomass than control biomass is just inoculated biomass/RGR
df.means$control_biomass <- df.means$plant_biomass_lsmean/df.means$RGR_lsmean

#Standardize trait and fitness values for each population separately
tmp <- df.means %>% group_by(Population) %>% summarize(
  pop_mean_tot_nod=mean(tot_nod_lsmean), 
  pop_sd_tot_nod=sd(tot_nod_lsmean),
  pop_mean_nod_mass=mean(nod_mass_lsmean),
  pop_sd_nod_mass=sd(nod_mass_lsmean),
  pop_mean_CHR=mean(CHR),
  pop_mean_SI=mean(SI),
  pop_mean_plant_biomass=mean(plant_biomass_lsmean),
  pop_sd_plant_biomass=sd(plant_biomass_lsmean),
  pop_mean_RGR=mean(RGR_lsmean),
  pop_sd_RGR=mean(RGR_lsmean)
  )

df.means <- merge(df.means, tmp, by="Population")

#Standardize traits
df.means$plant_biomass_std <- (df.means$plant_biomass_lsmean - df.means$pop_mean_plant_biomass)/df.means$pop_sd_plant_biomass
df.means$RGR <- (df.means$RGR_lsmean - df.means$pop_mean_RGR)/df.means$pop_sd_RGR

#Relativize fitness by dividing by the mean fitness
df.means$tot_nod_std <- df.means$tot_nod_lsmean/df.means$pop_mean_tot_nod
df.means$nod_mass_std <- df.means$nod_mass_lsmean/df.means$pop_mean_nod_mass
df.means$CHR_std <- df.means$CHR/df.means$pop_mean_CHR
df.means$SI_std <- df.means$SI/df.means$pop_mean_SI

#Total nodules and plant biomass

lm1 <- lm(tot_nod_std~plant_biomass_std, data=df.means) #Model
summary(lm1)
S1 <- paste0("S = ", round(summary(lm1)$coefficients[2,1], 3))
pval1 <- round(summary(lm1)$coefficients[2,4],3)
pval1 <- ifelse(pval1 <= 0.05, paste0("p = ", pval1), "ns")
S1 <- paste0(S1, ", ", pval1)

p1 <- ggplot(data=df.means, aes(y=tot_nod_std, x=plant_biomass_std, color=Population))+
      geom_point()+
      geom_smooth(method="lm", se=FALSE, color=1)+
      geom_text(aes(label=Strain),hjust=0, vjust=0, size=3)+
      ylab("Nodule number")+
      xlab("Plant biomass")+
      guides(color=FALSE)
p1

#Nodule mass and plant biomass

lm2 <- lm(nod_mass_std~plant_biomass_std, data=df.means) #Model
summary(lm2)
S2 <- paste0("S = ", round(summary(lm2)$coefficients[2,1], 3))
pval2 <- round(summary(lm2)$coefficients[2,4],3)
pval2 <- ifelse(pval2 <= 0.05, paste0("p = ", pval2), "ns")
S2 <- paste0(S2, ", ", pval2)

p2 <- ggplot(data=df.means, aes(y=nod_mass_std, x=plant_biomass_std, color=Population))+
      geom_point()+
      geom_smooth(method="lm", se=FALSE, color=1)+
      geom_text(aes(label=Strain),hjust=0, vjust=0, size=3)+
      ylab("Nodule mass")+
      xlab("Plant biomass")+
      guides(color=FALSE)
p2

#CHR and plant biomass

lm3 <- lm(CHR_std~plant_biomass_std, data=df.means) #Model
summary(lm3)
S3 <- paste0("S = ", round(summary(lm3)$coefficients[2,1], 3))
pval3 <- round(summary(lm3)$coefficients[2,4],3)
pval3 <- ifelse(pval3 <= 0.05, paste0("p = ", pval3), "ns")
S3 <- paste0(S3, ", ", pval3)

p3 <- ggplot(data=df.means, aes(y=CHR_std, x=plant_biomass_std, color=Population))+
      geom_point()+
      geom_smooth(method="lm", se=FALSE, color=1)+
      geom_text(aes(label=Strain),hjust=0, vjust=0, size=3)+
      ylab("CHR")+
      xlab("Plant biomass")+
      guides(color=FALSE)
p3

#SI and plant biomass

lm4 <- lm(SI_std~plant_biomass_std, data=df.means) #Model
summary(lm4)
S4 <- paste0("S = ", round(summary(lm4)$coefficients[2,1], 3))
p4 <- round(summary(lm4)$coefficients[2,4],3)
p4 <- ifelse(p4 <= 0.05, paste0("p = ", p4), "ns")
S4 <- paste0(S4, ", ", p4)

p4 <- ggplot(data=df.means, aes(y=SI_std, x=plant_biomass_std, color=Population))+
      geom_point()+
      geom_smooth(method="lm", se=FALSE, color=1)+
      geom_text(aes(label=Strain),hjust=0, vjust=0, size=3)+
      ylab("SI")+
      xlab("Plant biomass")+
      guides(color=FALSE)
p4

#Total nodules and RGR

lm5 <- lm(tot_nod_std~RGR, data=df.means) #Model
summary(lm5)
S5 <- paste0("S = ", round(summary(lm5)$coefficients[2,1], 3))
p5 <- round(summary(lm5)$coefficients[2,4],3)
p5 <- ifelse(p5 <= 0.05, paste0("p = ", p5), "ns")
S5 <- paste0(S5, ", ", p5)

p5 <- ggplot(data=df.means, aes(y=tot_nod_std, x=RGR, color=Population))+
      geom_point()+
      geom_smooth(method="lm", se=FALSE, color=1)+
      geom_text(aes(label=Strain),hjust=0, vjust=0, size=3)+
      ylab("Nodule number")+
      xlab("RGR")+
      guides(color=FALSE)
p5

#Nodule mass and RGR

lm6 <- lm(nod_mass_std~RGR, data=df.means) #Model
summary(lm6)
S6 <- paste0("S = ", round(summary(lm6)$coefficients[2,1], 3))
p6 <- round(summary(lm6)$coefficients[2,4],3)
p6 <- ifelse(p6 <= 0.05, paste0("p = ", p6), "ns")
S6 <- paste0(S6, ", ", p6)

p6 <- ggplot(data=df.means, aes(y=nod_mass_std, x=RGR, color=Population))+
      geom_point()+
      geom_smooth(method="lm", se=FALSE, color=1)+
      geom_text(aes(label=Strain),hjust=0, vjust=0, size=3)+
      ylab("Nodule number")+
      xlab("RGR")+
      guides(color=FALSE)
p6

#CHR and RGR

lm7 <- lm(CHR~RGR, data=df.means) #Model
summary(lm7)
S7 <- paste0("S = ", round(summary(lm7)$coefficients[2,1], 3))
p7 <- round(summary(lm7)$coefficients[2,4],3)
p7 <- ifelse(p7 <= 0.05, paste0("p = ", p7), "ns")
S7 <- paste0(S7, ", ", p7)

p7 <- ggplot(data=df.means, aes(y=CHR, x=RGR, color=Population))+
      geom_point()+
      geom_smooth(method="lm", se=FALSE, color=1)+
      geom_text(aes(label=Strain),hjust=0, vjust=0, size=3)+
      ylab("CHR")+
      xlab("RGR")+
      guides(color=FALSE)
p7

#SI and RGR

lm8 <- lm(SI~RGR, data=df.means) #Model
summary(lm8)
S8 <- paste0("S = ", round(summary(lm8)$coefficients[2,1], 3))
p8 <- round(summary(lm8)$coefficients[2,4],3)
p8 <- ifelse(p8 <= 0.05, paste0("p = ", p8), "ns")
S8 <- paste0(S8, ", ", p8)

p8 <- ggplot(data=df.means, aes(y=SI, x=RGR, color=Population))+
      geom_point()+
      geom_smooth(method="lm", se=FALSE, color=1)+
      geom_text(aes(label=Strain),hjust=0, vjust=0, size=3)+
      ylab("SI")+
      xlab("RGR")+
      guides(color=FALSE)
p8

plot_grid(p1,p2,p3,p4,p5,p6,p7,p8, nrow=2, ncol=4, labels=c(S1, S2, S3, S4, S5, S6, S7, S8), hjust=-1)

#Correlations among rhizobium fitness proxies
lm9 <- lm(SI~CHR, data=df.means) #Model
summary(lm9)
p9 <- round(summary(lm9)$coefficients[2,4],3)
r9 <- 
label9 <- expression("r"^2*" = ", r9)


p9 <- ggplot(data=df.means, aes(y=SI, x=CHR))+
      geom_point()+
      geom_smooth(method="lm")+
      xlab("CHR")+
      ylab("SI")+
      geom_text(aes(label=Strain),hjust=0, vjust=0)
p9

p10 <- ggplot(data=df.means, aes(y=SI, x=nod_mass_lsmean))+
      geom_point()+
      geom_smooth(method="lm")+
      geom_text(aes(label=Strain),hjust=0, vjust=0)
p10

p9 <- ggplot(data=df.means, aes(y=SI, x=tot_nod_lsmean))+
      geom_point()+
      geom_smooth(method="lm")+
      geom_text(aes(label=Strain),hjust=0, vjust=0)
p9


p10 <- ggplot(data=df.means, aes(y=CHR, x=nod_mass_lsmean))+
      geom_point()+
      geom_smooth(method="lm")+
      geom_text(aes(label=Strain),hjust=0, vjust=0)
p10

p11 <- ggplot(data=df.means, aes(y=CHR, x=tot_nod_lsmean))+
      geom_point()+
      geom_smooth(method="lm")+
      geom_text(aes(label=Strain),hjust=0, vjust=0)
p11

#Corrlations among plant fitness proxies

p12 <- ggplot(data=df.means, aes(y=RGR, x=plant_biomass_std))+
      geom_point()+
      geom_smooth(method="lm")+
      geom_text(aes(label=Strain),hjust=0, vjust=0)
p12




plot_grid(p9, label=p9)

p13 <- ggplot(data=df.means, aes(y=control_biomass, x=plant_biomass_lsmean))+
      geom_point()+
      geom_smooth(method="lm")+
      geom_text(aes(label=Strain),hjust=0, vjust=0)
p13

p13 <- ggplot(data=df.means, aes(y=control_biomass, x=SI))+
      geom_point()+
      geom_smooth(method="lm")+
      geom_text(aes(label=Strain),hjust=0, vjust=0)
p13

p13 <- ggplot(data=df.means, aes(y=control_biomass, x=CHR))+
      geom_point()+
      geom_smooth(method="lm")+
      geom_text(aes(label=Strain),hjust=0, vjust=0)
p13



lm <- lm(SI~plant_biomass_lsmean, data=df.means)
summary(lm)

lm <- lm(CHR~plant_biomass_lsmean, data=df.means)
summary(lm)

lm <- lm(CHR~plant_biomass_lsmean, data=df.means)
summary(lm)


lm <- lm(nod_mass_lsmean~plant_biomass_lsmean, data=df.means)
summary(lm)

```


```