---
title: "Are rhizobia under selection to cheat?"
author: "Megan Frederickson"
date: "`r format(Sys.Date())`"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Is there fitness conflict between legumes and rhizobia?

This repository re-analyzes the data in: Gano-Cohen KA, Wendlandt CE, Stokes PJ, Blanton MA, Quides KW, Zomorrodian A, Adinata ES, Sachs JL (2019) Interspecific conflict and the evolution of ineffective rhizobia. Ecology Letters. https://doi.org/10.1111/ele.13247

The authors make the case that in their legume-rhizobium study system, the rhizobia are selected to cheat. They base this conclusion on a negative correlation between legume and rhizobium fitnesses (their Figure 5).

Here, I re-analyze their data, mostly to explore how they measured rhizobium fitness.

```{r Load packages, message=FALSE, warning=FALSE, include=FALSE}

library(tidyverse) #includes ggplot2, dplyr, readr, stringr
library(knitr)
library(cowplot)
library(car)
library(glmm)
library(nlme)
library(lme4)
library(emmeans)

```

I downloaded their data from Dryad on April 16, 2019. The citation for the data package is: Gano-Cohen KA, Wendlandt CE, Stokes PJ, Blanton MA, Quides KW, Zomorrodian A, Adinata ES, Sachs JL (2019) Data from: Interspecific conflict and the evolution of ineffective rhizobia. Dryad Digital Repository. https://doi.org/10.5061/dryad.cr65269

First we need to read in the data, which is in three different tables in the Dryad package.

```{r Read in data, message=FALSE, warning=FALSE}

table_S4 <- read_csv("Table_S4.csv", col_types = cols(Strain = col_factor(levels = c("132", "133", "134", "135", "136", "137", "138", "139", "140", "141", "142", "143", "144", "145", "146", "147", "148", "149", "150", "151", "152", "153", "154", "155", "156", "157", "158", "159", "160", "161", "control")), `Host Line` = col_factor(levels = c("BMR01.03", "BMR07.03", "UnH: Cla12.04", "UnL: Anz13.04", "A. heermannii", "Gri01.01", "Gri01.13", "Cla10.01", "Cla01.04", "UCR02.07", "UCR09.05", "Yuc02.07", "Yuc02.01", "Anz11.01", "Anz10.01")), Block = col_factor(levels = c("1", "2", "3", "4", "5")), `CHR local abundance` = col_number(), `Mean individual  nodule biomass (mg)` = col_number(), `Mineral N (ppm)` = col_number(), `Plant #` = col_number(), Population = col_factor(levels = c("ANZ",  "BMR", "CLA", "GRI", "UCR", "YUC")), `Relative Growth` = col_number(), `Roots mass (g)` = col_number(), `Shoots mass (g)` = col_number(), `Total N (%)` = col_number(), `Total nodules` = col_number()))
  
table_S1 <- read_csv("Table_S1.csv", col_names = c("Full_Strain_Name", "Year", "Population", "glnII_Haplotype", "recA_Haplotype", "nodZ_Haplotype",  "nolL_Haplotype"), col_types = cols(Year = col_number(), Population = col_factor(levels = c("ANZ",  "BMR", "CLA", "GRI", "UCR", "YUC"))), skip = 1)

table_S2 <- read_csv("Table_S2.csv", col_names = c("Strain", "Full_Strain_Name", "Population", "Latitude", "Longitude", "glnII_Haplotype", "glnII_Accession", "recA_Haplotype", "recA_Accession", "nodZ_Haplotype", "nodZ_Accession", "nolL_Haplotype", "nolL_Accession", "CHR_Haplotype", "CHR genotype frequency" , "SI_haplotye", "SI genotype frequency"), col_types = cols(Strain = col_factor(levels = c("132", "133", "134", "135", "136", "137", "138", "139", "140", "141", "142", "143", "144", "145", "146", "147", "148", "149", "150", "151", "152", "153", "154", "155", "156", "157", "158", "159", "160", "161")), `CHR genotype frequency` = col_number(), `SI genotype frequency` = col_number(), Population = col_factor(levels = c("Bodega Marine Reserve", "Griffith Park", "Robert J. Bernard Biological Field Station", "University of California Riverside", "Burns Pinon Ridge Reserve", "Anza Borrego Desert State Park"))), skip = 2)

```

Next, we need to clean up the data and wrangle it into a single dataframe.

```{r Wrangle data, message=FALSE, warning=FALSE}

#There is an issue with the "Relative Growth" column in the Dryad version of Table S4
#Values are hundreds or thousands, but inoculated plants did not grow 100-1000x more than controls
#In fact, the x-axis of Figure 5 in the original paper shows values from ~0.5-1.2 on a log scale
#So real values should be on the order of ~1-100x
#As a result, I recalculated Relative Growth Rate (RGR) from scratch
#I followed the methods described in the original paper, attempting to match the original analysis
table_S4$pop_block <- paste0(table_S4$Block, table_S4$`Host Line`) #Make a unique identifier for each block-host line combination
table_S4$Plant_biomass <- table_S4$`Shoots mass (g)`+table_S4$`Roots mass (g)` #Sum root and shoot mass
tmp <- subset(table_S4, Strain == "control") #Subset control plants only
table_S4 <- merge(table_S4, tmp[, c(1, 17, 18)], by="pop_block") #Match inoculated plants to controls based on which block and host line they were

#The footnotes for Table 1 in the original paper says when controls were substituted, so I follow suit here
table_S4[table_S4$`Plant #.y` == 111, 20] <- table_S4[table_S4$`Plant #.y` == 106, 20] #Sub 106 for 111
table_S4[table_S4$`Plant #.y` == 222, 20] <- table_S4[table_S4$`Plant #.y` == 216, 20] #Sub 216 for 222
table_S4[table_S4$`Plant #.y` == 669, 20] <- mean(c(unique(table_S4[table_S4$`Plant #.y` == 651, 20]), unique(table_S4[table_S4$`Plant #.y` == 660, 20]), unique(table_S4[table_S4$`Plant #.y` == 663, 20]), unique(table_S4[table_S4$`Plant #.y` == 680, 20]), unique(table_S4[table_S4$`Plant #.y` == 684, 20]))) #Sub mean of 651, 660, 663, 680, and 684 for 669

table_S4 <- subset(table_S4, `Shoots mass (g)` != "DEAD") #Exclude dead plants, as per paper
table_S4 <- subset(table_S4, Strain != "control") #Exclude uninoculated controls, as per paper
table_S4 <- subset(table_S4, `Total nodules` > 0) #Exclude inoculated plants that formed no nodules, as per paper
table_S4$RGR <- table_S4$Plant_biomass.x/table_S4$Plant_biomass.y #Re-calculate RGR (values look better)
table_S4$logRGR <- log10(table_S4$RGR) #Log-transform RGR, as per paper

table_S1$SI_haplotype <- paste0(table_S1$nodZ_Haplotype, "_", table_S1$nolL_Haplotype) #Concatenate SI haplotypes, as per paper
table_S1$CHR_haplotype <- paste0(table_S1$glnII_Haplotype, "_", table_S1$recA_Haplotype) #Concatenate CHR haplotypes, as per paper

table_S2$Population <- ifelse(table_S2$Population == "Bodega Marine Reserve", "BMR", ifelse(table_S2$Population == "Griffith Park", "GRI", ifelse(table_S2$Population == "Robert J. Bernard Biological Field Station", "CLA", ifelse(table_S2$Population == "University of California Riverside", "UCR", ifelse(table_S2$Population == "Anza Borrego Desert State Park", "ANZ", "YUC"))))) #Abbreviate Table_S2 population names, to make them match across data tables

table_S1$Plant_ID <- gsub('R.*', "", (toupper(gsub('_.*', "", table_S1$Full_Strain_Name)))) #Make a column of unique plant ids

table_S1 <- table_S1[,c(1, 10, 2:9)] #Reorder columns
table_S1.long <- gather(table_S1, locus, haplotype, glnII_Haplotype:CHR_haplotype, factor_key=TRUE) #Make wide data into long format
table_S1.long <- subset(subset(table_S1.long, haplotype != "n/an/a"), haplotype != "n/a_n/a")  #Remove NAs
table_S1.long <- subset(table_S1.long, locus == "SI_haplotype" | locus == "CHR_haplotype") #Subset to just CHR and SI haplotypes

#Calculate nodules and plants sampled per population
new.table <- table_S1.long %>% group_by(Population, locus) %>% summarize(total_nods_sampled=n(), total_plants_sampled=length(unique(Plant_ID))) #Summarize data by locus and population
new.table.long <- merge(subset(new.table, locus == "SI_haplotype"), subset(new.table, locus == "CHR_haplotype"), by="Population") #Make wide data long
new.table.long <- new.table.long[, c(1,3,4,6,7)]
colnames(new.table.long) <- c("Population", "SI_nods_sampled", "SI_plants_sampled", "CHR_nods_sampled", "CHR_plants_sampled")

#Merge number of nodules and plants sampled with Table S2 data
table_S2 <- merge(table_S2, new.table.long, by="Population", all.x = TRUE) #Add sampling effort to Table 2

#Calculate strain means in Table S4
#These should really be predicted from a linear mixed model using something like the emmeans package
#But the original paper says they just used raw averages, so I do so here too
#Also, because the paper calculated strain means just on sympatric host lines, I do the same here
data_sym <- subset(table_S4, `Host Line` != "A. heermannii" & `Host Line` != "UnH: Cla12.04" & `Host Line` != "UnL: Anz13.04")
new.table2 <- data_sym %>% group_by(Population, Strain) %>% summarize(mean_RGR = mean(`Relative Growth`, na.rm=TRUE), mean_total_nodules = mean(`Total nodules`, na.rm=TRUE), mean_nodule_mass = mean(`Mean individual  nodule biomass (mg)`, na.rm=TRUE), mean_log10_RGR = mean(logRGR, na.rm=TRUE)) #Calculate means

#Merge data in Tables S2 and S4 into a single data frame
df <- merge(table_S2, new.table2[ ,2:6], by="Strain") 

```

## How many nodules and plants were sampled per site?

```{r Sampling, message=FALSE, warning=FALSE}
colnames(new.table) <- c("Population", "Locus", "Nodules sampled (no.)", "Plants sampled (no.)")
new.table$Locus <- gsub("_haplotype", "", new.table$Locus)
kable(new.table)
```

The sampling is very uneven, with few plants sampled in ANZ and YUC. SI genotype frequencies are calculated from just 2 plants at YUC, and CHR genotype frequencies are calculated from just 4 plants at ANZ and 7 plants at YUC.

## Plot sampling of CHR and SI frequencies 

Does this uneven sampling matter? Here, I plot the relationship between genotype frequency at a site and sampling effort (i.e., the number of plants sampled). I also use simple linear models to ask whether genotype frequencies are related to sampling effort.

```{r Data distributions, message=FALSE, warning=FALSE, dpi=600, fig.width=8, fig.height=6}

#Model relationship between CHR frequency and number of plants sampled
model1 <- lm(`CHR genotype frequency`~CHR_plants_sampled, data=df)
summary(model1) #Significant correlation 

#Model relationship between SI frequency and number of plants sampled
model2 <- lm(`SI genotype frequency`~SI_plants_sampled, data=df)
summary(model2) #Significant correlation

#Make figures
CHR <- ggplot(data=df, aes(y=`CHR genotype frequency`, x=CHR_plants_sampled, color=Population))+geom_point()+ geom_smooth(method="lm", se=TRUE, color=1)+xlab("Plants sampled (no.)")+ylab("CHR genotype frequency")+geom_text(aes(label=Strain),hjust=0, vjust=0, size=2.5, nudge_x = 0.05, check_overlap=TRUE)+scale_x_continuous(limits=c(0, 35))+theme(legend.position = c(0.7,0.7))

SI <- ggplot(data=df, aes(y=`SI genotype frequency`, x=SI_plants_sampled, color=Population)) +geom_smooth(method="lm", se=TRUE, color=1)+geom_point()+xlab("Plants sampled (no.)")+ylab("SI genotype frequency")+geom_text(aes(label=Strain),hjust=0, vjust=0, size=2.5, nudge_x = 0.05, check_overlap=TRUE)+scale_x_continuous(limits=c(0, 23))+guides(color=FALSE)

fig <- plot_grid(CHR, SI, nrow=2, labels="auto") 
fig #Show figure
```

## Rarefaction analysis of haplotypes sampled at field sistes

Here, I use standard methods from community ecology to explore whether rhizobia genotypes were adequately sampled at the six field sites.  

```{r Rarefaction, message=FALSE, warning=FALSE, dpi=600, fig.width=8, fig.height=6}

library(vegan) #Load libaries with rarefaction functions
library(labdsv)

table_S1$pop_plant <- paste0(table_S1$Population, table_S1$Plant_ID) #Create unique population-plant identifier
  
#Transform Table S1 data into a community matrix, in which columns are SI or CHR haplotypes and rows are plants
SI_matrix <- table_S1[, c(9,11)] %>% group_by(pop_plant, SI_haplotype) %>% count(SI_haplotype)
SI_matrix <- subset(SI_matrix, SI_haplotype != "n/a_n/a") #Remove NAs
SI_matrix <- subset(SI_matrix, SI_haplotype != "Z01_X" & SI_haplotype != "Z13_X") #Remove Xs 
SI_matrix <- matrify(as.data.frame(SI_matrix))
CHR_matrix <- table_S1[, c(10,11)] %>% group_by(pop_plant, CHR_haplotype) %>% count(CHR_haplotype)
CHR_matrix <- subset(CHR_matrix, CHR_haplotype != "n/a_n/a") #Remove NAs
CHR_matrix <- matrify(as.data.frame(CHR_matrix))

#Rarefy within each population
SI.accum.CLA <- specaccum(SI_matrix[21:40, ])
SI.accum.BMR <- specaccum(SI_matrix[5:20, ])
SI.accum.ANZ <- specaccum(SI_matrix[1:4, ]) 
SI.accum.GRI <- specaccum(SI_matrix[41:44, ])
SI.accum.YUC <- specaccum(SI_matrix[50:51, ])
SI.accum.UCR <- specaccum(SI_matrix[45:49, ])
CHR.accum.CLA <- specaccum(CHR_matrix[21:40, ])
CHR.accum.BMR <- specaccum(CHR_matrix[5:20, ])
CHR.accum.ANZ <- specaccum(CHR_matrix[1:4, ]) 
CHR.accum.GRI <- specaccum(CHR_matrix[41:58, ])
CHR.accum.YUC <- specaccum(CHR_matrix[90:96, ])
CHR.accum.UCR <- specaccum(CHR_matrix[59:89, ])

#Wrangle in to a single dataframe
SI.accum <- rbind(data.frame(plants=SI.accum.YUC$sites, richness=SI.accum.YUC$richness, SD=SI.accum.YUC$sd, Population="YUC"), data.frame(plants=SI.accum.UCR$sites, richness=SI.accum.UCR$richness, SD=SI.accum.UCR$sd, Population="UCR"), data.frame(plants=SI.accum.GRI$sites, richness=SI.accum.GRI$richness, SD=SI.accum.GRI$sd, Population="GRI"), data.frame(plants=SI.accum.ANZ$sites, richness=SI.accum.ANZ$richness, SD=SI.accum.ANZ$sd, Population="ANZ"), data.frame(plants=SI.accum.BMR$sites, richness=SI.accum.BMR$richness, SD=SI.accum.BMR$sd, Population="BMR"), data.frame(plants=SI.accum.CLA$sites, richness=SI.accum.CLA$richness, SD=SI.accum.CLA$sd, Population="CLA"))
SI.accum$Population <- factor(SI.accum$Population, levels=c("ANZ", "BMR", "CLA", "GRI", "UCR", "YUC"))

CHR.accum <- rbind(data.frame(plants=CHR.accum.YUC$sites, richness=CHR.accum.YUC$richness, SD=CHR.accum.YUC$sd, Population="YUC"), data.frame(plants=CHR.accum.UCR$sites, richness=CHR.accum.UCR$richness, SD=CHR.accum.UCR$sd, Population="UCR"), data.frame(plants=CHR.accum.GRI$sites, richness=CHR.accum.GRI$richness, SD=CHR.accum.GRI$sd, Population="GRI"), data.frame(plants=CHR.accum.ANZ$sites, richness=CHR.accum.ANZ$richness, SD=CHR.accum.ANZ$sd, Population="ANZ"), data.frame(plants=CHR.accum.BMR$sites, richness=CHR.accum.BMR$richness, SD=CHR.accum.BMR$sd, Population="BMR"), data.frame(plants=CHR.accum.CLA$sites, richness=CHR.accum.CLA$richness, SD=CHR.accum.CLA$sd, Population="CLA"))
CHR.accum$Population <- factor(CHR.accum$Population, levels=c("ANZ", "BMR", "CLA", "GRI", "UCR", "YUC"))

#Plot 'genotype accumulation' curves
SI.accum.curve <- ggplot(data=SI.accum, aes(x=plants, y=richness, color=Population))+geom_point()+geom_line()+geom_errorbar(aes(x=plants, ymin=richness-SD, ymax=richness+SD), alpha=0.5, width=0.1)+xlab("Plants sampled (no.)")+ylab("Unique SI genotypes (no.)")+scale_x_continuous(limits=c(0, 23))+guides(color=FALSE)

CHR.accum.curve <- ggplot(data=CHR.accum, aes(x=plants, y=richness, color=Population))+geom_point()+geom_line()+geom_errorbar(aes(x=plants, ymin=richness-SD, ymax=richness+SD), alpha=0.5, width=0.1)+xlab("Plants sampled (no.)")+ylab("Unique CHR genotypes (no.)")+scale_x_continuous(limits=c(0, 35))+guides(color=FALSE)
 
#Combine with previous figures to make full figure for paper
fig1 <- plot_grid(CHR, SI, CHR.accum.curve, SI.accum.curve, nrow=2, ncol=2, labels="auto")
fig1
save_plot("Fig1.png", fig1, base_width=8, base_height=8)
```

## How does uneven sampling among sites affect the results?

I corrected for uneven sampling in two ways: 1) by relativizing fitness by dividing by the population mean, and 2) by sub-sampling the data for each population to the smallest sample sizes, and re-calculating genotype frequencies. Then I re-analyzed the relationship between strain frequency and symbiotic effectiveness (i.e., what appears in the original paper's Figure 5) using simple linear models. 

## Method 1: Relativize fitness within populations

First, I need to re-create the original analysis in the paper's Figure 5, to make sure I get the same answer.

```{r Fig5, message=FALSE, warning=FALSE, dpi=600, fig.width=8, fig.height=6}

#Model CHR genotype frequency, as in original paper
model3 <- lm(`CHR genotype frequency`~mean_log10_RGR, data=df)
summary(model3) #Numbers match paper!
Anova(model3) #Numbers match paper!

#Model SI genotype frequency, as in original paper
model4 <- lm(`SI genotype frequency`~mean_log10_RGR, data=df)
summary(model4) #Numbers match paper
Anova(model4) #Numbers match paper

#Make my own version of the authors' Figure 5
orig.fig5.CHR <- ggplot(data=df, aes(y=`CHR genotype frequency`, x=mean_log10_RGR, color=Population))+geom_smooth(method="lm", color=1)+geom_point()+xlab("Symbiotic effectiveness")+ylab("CHR genotype frequency")+geom_text(aes(label=Strain),hjust=0, vjust=0, size=2.5, check_overlap=TRUE)+theme(legend.position = c(0.7, 0.7))

orig.fig5.SI <- ggplot(data=df, aes(y=`SI genotype frequency`, x=mean_log10_RGR, color=Population)) +geom_smooth(method="lm", color=1)+geom_point()+xlab("Symbiotic effectiveness")+ylab("SI genotype frequency")+ geom_text(aes(label=Strain),hjust=0, vjust=0, size=2.5,  check_overlap=TRUE)+guides(color=FALSE)

orig.fig5 <-plot_grid(orig.fig5.CHR, orig.fig5.SI, nrow=2, labels="auto")
orig.fig5 #Hmm ... very close, but not exactly the same

```

## Relative fitness within populations 

Ideally, to compare across studies, fitness measures, and traits, we should calculate selection gradients in the standard way, as we would for any continuous phenotype. Normally, fitness is relativized by dividing by population mean fitness, and traits are standardized by subtracting the mean and dividing by the standard deviation. This allows comparisons of the strength of selection across analyses because everything is on a common scale. 

Here, I first relativize fitness within populations by dividing by population mean fitness (i.e., population mean strain frequency). Then I use the same strain means for relative growth as above in one analysis, and then also  standardize strain means by substracting the mean and dividing by the SD in a second analysis. Either way, the relationship between genotype frequency and symbiotic effectiveness is non-significant (for both CHR and SI) once fitness is relativized within each population. The only difference that standardizing the trait data makes is to generate an estimate of the strength of selection (i.e., beta) that is interpretable relative to other studies. It does not change R-squared or p values. 

```{r Relative fitness and standardize traits, message=FALSE, warning=FALSE, dpi=600, fig.width=8, fig.height=6}

tmp <- df %>% group_by(Population) %>% summarize(pop_mean_CHR=mean(as.numeric(`CHR genotype frequency`), na.rm=TRUE), pop_mean_SI=mean(as.numeric(`SI genotype frequency`), na.rm=TRUE)) #Calculate mean fitness for each population
df <- merge(df, tmp, by="Population") #Merge data frames
df$CHR_std <- df$`CHR genotype frequency`/df$pop_mean_CHR #Relative fitness within each population
df$SI_std <- df$`SI genotype frequency`/df$pop_mean_SI #Relative fitness within each population
df$RGR_std <- (df$mean_log10_RGR - mean(df$mean_log10_RGR, na.rm=TRUE))/sd(df$mean_log10_RGR, na.rm=TRUE) #Standardize trait by subtracting the mean and dividing by the standard deviation

#Model relativized CHR and unstandardized RGR
model5 <- lm(CHR_std~mean_log10_RGR, data=df)
summary(model5) #Non-significant

#Model relativized CHR and standardized RGR
model6 <- lm(CHR_std~RGR_std, data=df)
summary(model6) #Non-significant

#Model relativized SI and unstandardized RGR
model7 <- lm(SI_std~mean_log10_RGR, data=df) 
summary(model7) #Non-significant

#Relativized SI and standardized RGR
model8 <- lm(SI_std~RGR_std, data=df) 
summary(model8) #Non-significant

#Make new versions of Figure 5
new.Fig5.CHR.unstandardized <- ggplot(data=df, aes(y=CHR_std, x=mean_log10_RGR, color=Population))+ geom_point()+geom_smooth(method="lm", se=TRUE, linetype="dotted", color="grey")+geom_text(aes(label=Strain),hjust=0, vjust=0, size=3, check_overlap=TRUE)+ylab("CHR genotype frequency")+xlab("Symbiotic effectiveness")+guides(color=FALSE)

new.Fig5.SI.unstandardized <- ggplot(data=df, aes(y=SI_std, x=mean_log10_RGR, color=Population))+geom_point()+geom_smooth(method="lm", se=TRUE, linetype="dotted", color="grey")+geom_text(aes(label=Strain),hjust=0, vjust=0, size=3, check_overlap=TRUE)+ ylab("SI genotype frequency")+ xlab("Symbiotic effectiveness")+guides(color=FALSE)

new.Fig5ab <- plot_grid(new.Fig5.CHR.unstandardized, new.Fig5.SI.unstandardized, nrow=2, labels="auto")
new.Fig5ab

```

## Method 2: Down-sample each population to minimum sample sizes and re-calculate genotype frequencies

Finally, I down-sampled each population 100 times to either 2 or 4 plants for SI and CHR, respectively, as these were the minimum number of plants sampled per population to calculate SI or CHR frequencies, and then I recalculated genotype frequencies. For each population, I took the mean genotype frequency across the 100 iterations and re-analyzed the relationship between genotype frequency and symbiotic effectiveness. Again, the relationships were never significatnt after accounting for uneven sampling. 

```{r Downsample, message=FALSE, warning=FALSE, dpi=600, fig.width=8, fig.height=6}

#I did this separately for CHR and SI genotype frequencies
#CHR first
CHR.min <- 4 #Minimum number of plants sampled (in ANZ population)
pop <- c("UCR", "CLA", "GRI", "BMR", "YUC", "ANZ") #Create vector of population names
df.CHR <- data.frame(Population=character(), haplotype = character(), n=double(), tot_n=double(), freq=double(), stringsAsFactors=FALSE) #Initialize empty frame to store all the sub-sampled data in

#Two for loops that loop through 100 iterations for each population
for (i in 1:6) {
  tmp.pop <- pop[i]
for (j in 1:500) {
  tmp <- subset(table_S1, Population == tmp.pop)
  tmp.plants <- sample(unique(tmp$Plant_ID), CHR.min, replace = FALSE)
  tmp.data <- subset(tmp, tmp$Plant_ID %in% tmp.plants)
  tmp.data.long <- gather(tmp.data, locus, haplotype, glnII_Haplotype:CHR_haplotype, factor_key=TRUE)
  tmp.data.long <- subset(tmp.data.long, locus == "CHR_haplotype")
  tmp.data.long <- subset(subset(tmp.data.long, haplotype != "n/an/a"), haplotype != "n/a_n/a")
  tmp2 <- tmp.data.long %>% group_by(Population, haplotype) %>% summarize(n=n())
  tmp2$tot_n <- length(tmp.data.long$haplotype)
  tmp2$freq <- tmp2$n/tmp2$tot_n
  df.CHR<-rbind(df.CHR, as.data.frame(tmp2))
  }
}

#Merge re-sampled and original data by CHR haplotype and population
df$popCHR <- paste0(df$Population, df$CHR_Haplotype)
df.CHR$popCHR <- paste0(df.CHR$Population, df.CHR$haplotype)
df.CHR <- merge(df.CHR, df, by="popCHR")

#Calculate mean genotype frequencies per strain from 100 iterations
CHR.sum <- df.CHR %>% group_by(Population.x, haplotype) %>% summarize(n=n(), mean_CHR_freq=mean(freq, na.rm=TRUE), sd=sd(freq, na.rm=TRUE), se=sd/sqrt(n))
CHR.sum$popCHR <- paste0(CHR.sum$Population.x, CHR.sum$haplotype)
CHR.sum <- merge(CHR.sum, df, by="popCHR")

#Now SI 
SI.min <- 2 #Minimum number of plants sampled (in YUC population)
df.SI <- data.frame(Population=character(), haplotype = character(), n=double(), tot_n=double(), freq=double(), stringsAsFactors=FALSE) #Initialize empty frame to store all the sub-sampled data in

#Again, two for loops that loop through 100 iterations for each population
for (i in 1:6) {
  tmp.pop <- pop[i]
  for (j in 1:100) {
  tmp <- subset(table_S1, Population == tmp.pop & SI_haplotype != "n/a_n/a")
  tmp.plants <- sample(unique(tmp$Plant_ID), SI.min, replace = FALSE)
  tmp.data <- subset(tmp, tmp$Plant_ID %in% tmp.plants)
  tmp.data.long <- gather(tmp.data, locus, haplotype, glnII_Haplotype:CHR_haplotype, factor_key=TRUE)
  tmp.data.long <- subset(tmp.data.long, locus == "SI_haplotype")
  tmp.data.long <- subset(subset(tmp.data.long, haplotype != "n/an/a"), haplotype != "n/a_n/a")
  tmp2 <- tmp.data.long %>% group_by(Population, haplotype) %>% summarize(n=n())
  tmp2$tot_n <- length(tmp.data.long$haplotype)
  tmp2$freq <- tmp2$n/tmp2$tot_n
  df.SI<-rbind(df.SI, as.data.frame(tmp2))
  }
}
  
#Merge re-sampled and original data by SI haplotype and population
df$popSI <- paste0(df$Population, df$SI_haplotye)
df.SI$popSI <- paste0(df.SI$Population, df.SI$haplotype)
df.SI <- merge(df.SI, df, by="popSI")

#Calculate mean genotype frequencies per strain from 100 iterations
SI.sum <- df.SI %>% group_by(Population.x, haplotype) %>% summarize(n=n(), mean_SI_freq=mean(freq, na.rm=TRUE), sd=sd(freq, na.rm=TRUE), se=sd/sqrt(n))
SI.sum$popSI <- paste0(SI.sum$Population.x, SI.sum$haplotype)
SI.sum <- merge(SI.sum, df, by="popSI")

#Fit model for mean CHR genotype frequency from re-sampling and RGR
model9 <- lm(mean_CHR_freq~mean_log10_RGR, data=CHR.sum)
summary(model9) #Non-significant

#Fit model for mean SI genotype frequency from re-sampling and RGR
model10 <- lm(mean_SI_freq~mean_log10_RGR, data=SI.sum)
summary(model10) #Non-significant

#Make figures
new.Fig5.CHR.downsampled <-ggplot(data=CHR.sum, aes(x=mean_log10_RGR, y=mean_CHR_freq, color=Population.x))+
geom_point()+geom_smooth(method="lm", linetype="dotted", color="grey")+geom_text(aes(label=Strain),hjust=0, vjust=0, size=3, check_overlap=TRUE)+ylab("CHR genotype frequency")+xlab("Symbiotic effectiveness")+guides(color=FALSE)

new.Fig5.SI.downsampled <-ggplot(data=SI.sum, aes(x=mean_log10_RGR, y=mean_SI_freq, color=Population.x))+
geom_point()+geom_smooth(method="lm", linetype="dotted", color="grey")+geom_text(aes(label=Strain),hjust=0, vjust=0, size=3, check_overlap=TRUE)+ylab("SI genotype frequency")+xlab("Symbiotic effectiveness")+guides(color=FALSE)

full.new.Fig5 <- plot_grid(new.Fig5.CHR.unstandardized, new.Fig5.SI.unstandardized, new.Fig5.CHR.downsampled, new.Fig5.SI.downsampled, nrow=2, ncol=2, labels="auto")
full.new.Fig5
save_plot("Fig2.png", full.new.Fig5, base_width=8, base_height=8)
  
```


